---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# vegadown

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/vegadown)](https://CRAN.R-project.org/package=vegadown)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![R-CMD-check](https://github.com/vegawidget/vegadown/workflows/R-CMD-check/badge.svg)](https://github.com/vegawidget/vegadown/actions)
<!-- badges: end -->

The goal of vegadown is to help you compose Vega and Vega-Lite specs in RMardown documents.

## Installation

This package is not (yet) on CRAN. 
You can install the dev version of vegadown from GitHub with:

```r
# install.packages("devtools")
devtools::install_github("vegawidget/vegadown")
```

## Example

```{r setup}
library("vegadown")
```

## Code of Conduct

Please note that the vegadown project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
