---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  dev = "svg"
)
```

# vegadown

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/vegadown)](https://CRAN.R-project.org/package=vegadown)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![R-CMD-check](https://github.com/vegawidget/vegadown/workflows/R-CMD-check/badge.svg)](https://github.com/vegawidget/vegadown/actions)
<!-- badges: end -->

The goal of vegadown is to help you compose Vega and Vega-Lite specs within RMarkdown documents.

The audience for this package are people who are already comfortable with Vega-Lite JSON specs. 
Compared with using the [vegawiget](https://vegawidget.github.io/vegawidget) package alone, vegadown offers a quicker, more minimal, environment for developing specs.
However, it still does not provide humane feedback when something goes wrong.  

## Installation

This package is not (yet) on CRAN. 
You can install the development version from GitHub with:

```r
# install.packages("devtools")
devtools::install_github("vegawidget/vegadown")
```

## Example

```{r setup}
library("vegadown")
library("palmerpenguins")
```

Using vegawidget alone, it can be a bit awkward to compose specs using lists. 

Using this package, you can compose using JSON or YAML within RMarkdown chunks by specifying the chunk's language as `vegajson` or `vegayaml` (`vegayml`).
You can interpolate objects from your R environment by using the delimiters `${}`, much like you would using the [glue](https://glue.tidyverse.org/) package.

Here's what your code chunk would look like:

````
```{vegajson json-cars}`r ''`
{
  "$schema": "${vega_schema()}",
  "data": {"values": "${penguins}"},
  "mark": "point",
  "encoding": {
    "x": {
      "field": "bill_length_mm",
      "type": "quantitative",
      "scale": {"zero": false}
    },
    "y": {
      "field": "bill_depth_mm",
      "type": "quantitative",
      "scale": {"zero": false}
    },
    "color": {"field": "species", "type": "nominal"}
  }
}
```
````

Here's how it would appear:

```{vegajson json-penguins, echo=FALSE}
{
  "$schema": "${vega_schema()}",
  "data": {"values": "${penguins}"},
  "mark": "point",
  "encoding": {
    "x": {
      "field": "bill_length_mm",
      "type": "quantitative",
      "scale": {"zero": false}
    },
    "y": {
      "field": "bill_depth_mm",
      "type": "quantitative",
      "scale": {"zero": false}
    },
    "color": {"field": "species", "type": "nominal"}
  }
}
```

You can access your vegaspecs elsewhere in your RMarkdown document, using the `vegadown()` function with the chunk label:

```{r}
str(vegadown("json-penguins"))
```

For more details, please see the [getting started article](https://vegawidget.github.io/vegadown/articles/vegadown.html), which also shows:

- how to compose a vegaspec using YAML.
- more on interpolating objects from your R environment into `vegajson` or `vegayaml` (`vegayml`).

## Code of Conduct

Please note that the vegadown project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
