---
title: "vegadown"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vegadown}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library("vegadown")

# makes it easier to set data URL in specs
vw_set_base_url("https://cdn.jsdelivr.net/npm/vega-datasets@2")
```

Using only [vegawidget](https://vegawidget.github.io/vegawidget), you have to compose a Vega or Vega-Lite spec (vegaspec) using lists:

```{r vwcars}
as_vegaspec(
  list(
    `$schema` = vega_schema(),
    data = list("url" = "data/cars.json"),
    mark = "point",
    encoding = list(
      x = list(field = "Horsepower", type = "quantitative"),
      y = list(field = "Miles_per_Gallon", type = "quantitative")
    )  
  ) 
)
```

It can feel a bit awkward to code using so many lists. 

## First steps

Using vegadown, you can compose specs using JSON or YAML code chunks in RMarkdown:

- use `vegajson` or `vegayaml` (`vegayml`) as the language of your code chunk.
- afterwards, you can access the vegaspec of a given chunk using `vegadown()`.

If you label a code chunk as `json-cars`, the vegaspec will be available using `vegadown("json-cars")`.
You can get the names of available specs using `vegadown_names()`.

### Examples

Here's what your code chunk would look like:

````
```{vegajson json-cars}`r ''`
{
  "$schema": "https://vega.github.io/schema/vega-lite/v4.json",
  "data": {"url": "data/cars.json"},
  "mark": "point",
  "encoding": {
    "x": {"field": "Horsepower", "type": "quantitative"},
    "y": {"field": "Miles_per_Gallon", "type": "quantitative"}
  }
}
```
````

Here's how it would appear:

```{vegajson json-cars, echo=FALSE}
{
  "$schema": "https://vega.github.io/schema/vega-lite/v4.json",
  "data": {"url": "data/cars.json"},
  "mark": "point",
  "encoding": {
    "x": {"field": "Horsepower", "type": "quantitative"},
    "y": {"field": "Miles_per_Gallon", "type": "quantitative"}
  }
}
```

As noted above, you can also use YAML (`vegayaml` or `vegayml`); we can also specify sizing options, just as with vegawidget. Here's what the code-chunk front matter would look like:

````
```{vegayaml yaml-cars, vega.height=300, vega.width=300}`r ''`
```
````

Here's how it would look in your rendered document:

```{vegayaml yaml-cars, vega.height=300, vega.width=300}
$schema: "https://vega.github.io/schema/vega-lite/v4.json"
data: 
  url: "data/cars.json"
mark: point
encoding: 
  x: 
    field: Weight_in_lbs
    type: quantitative
  "y": 
    field: Miles_per_Gallon
    type: quantitative
```

Note that we quoted the `"y"` for the y-encoding; this is because of the (many) [reserved Boolean keywords](https://yaml.org/type/bool.html) in YAML.

You can access a particular spec using `vegadown()`:

```{r}
vegadown("json-cars")
```

To get the labels of chunks that use vegadown (thus far in the RMarkdown document):

```{r}
vegadown_labels()
```

## Interpolation from R

Writing a spec by hand is feasible if your data is available through a URL.
However, it would be tedious to specify the data values manually.
This is where interpolation is useful.

Let's introduce this topic using [palmerpenguins](https://allisonhorst.github.io/palmerpenguins/index.html), by Allison Horst, Alison Hill, and Kristen Gorman:

```{r}
library("palmerpenguins")

penguins
```

Let's also choose an adjective for the title of our chart:

```{r}
adjective <- sample(
  c("wonderful", "amazing", "interesting", "penguiny"),
  size = 1
)
```

We use [glue](https://glue.tidyverse.org/)-like notation to interpolate objects from the R environment into the spec; in fact, behind the scenes, we use `glue::glue()`.
Instead of using `{}` as interpolation delimiters, we use the JavaScript interpolation delimiters, `${}`:

```{vegayaml interp}
$schema: "${vega_schema()}"
data: 
  values: ${penguins}
title: "Such ${adjective} penguins!"
mark: point
encoding: 
  x: 
    field: bill_length_mm
    type: quantitative
    scale:
      zero: false
  "y": 
    field: bill_depth_mm
    type: quantitative
    scale:
      zero: false
  color:
    field: species
    type: nominal
```

For YAML, these delimiters work either quoted or unquoted. 

There are two interpolation contexts: 

- **substitution**, e.g. `"${vega_schema()}"`, `"${penguins}"`: if the entire string is interpolated, it substitutes the object itself; the object is not coerced to a string.
  This lets us interpolate data frames into a spec.
- **inline**, e.g. `"Such ${adjective} penguins!"`: this is traditional string-interpolation; the interpolated object is coerced to a string. 

The structure of the spec shows the interpolation:

```{r}
str(vegadown("interp"))
```

## Behind the scenes

There's not a lot to this package; it offers:

- re-exported [vegawidget](https://vegawidget.github.io/vegawidget/) functions to handle the rendering.
- functions to access the parsed specs: `vegadown()`, `vegadown_labels()`.
- some internal functions to handle interpolation from the R environment.
- [custom-language knitr engines](https://bookdown.org/yihui/rmarkdown-cookbook/custom-engine.html) for `vegajson` and `vegayaml` (`vegayml`).

Putting this together, I learned a lot by studying the [knitr-engine code for r2d3](https://github.com/rstudio/r2d3/blob/master/R/knit.R). 
As well, this [blog post](https://trestletech.com/2013/04/package-wide-variablescache-in-r-package/) by Jeff Allen had exactly what I needed to learn to get the `vegadown()` function to work.

